MSDS6372Project1

### 2021-01-26 MSDS6372Project1.Rmd
deleted MSRP over 1mil updated models accordingly...

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
# You need to adjust the work environment to run this code
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
# Directly import cleaned dataset and start from here
CarData = read.csv("CleanedCarData.csv")
# Convert factor variables
variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
summary(CarData)
# Plot agsint numeric variables
# CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
CarData = CarData[,-10] #delete MarketCategory
# Try different way of sampling by Models

######### Try Something else to split data ############
# First 842 are Models that only have 5 observations or less
ModelList = CarData%>% group_by(Model) %>% mutate(ModelCount = n()) %>% arrange(ModelCount)
#arrange(ModelList,desc(count))%>%print(n = Inf)
# Can only split train/test/validate set on Models that has 6 observations or more
ModelList = ModelList[,-27]
Not.Able.To.Split = ModelList[(1:842),]
Able.To.Split = ModelList[-(1:842),]

@@

# Joining train set with dataset that been left out
train = bind_rows(train,Not.Able.To.Split)
dim(train)

str(train)
str(CarData)
# Use Forward variable selection method to select possible variables as predictors
########################################
# Try Forward selection model.##########
########################################
library(leaps)
reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=966)
par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.fwd,305)
# By looking at this model, it has selected following predictors:
# Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback
forward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+VehicleStyle+Crossover+HighPerformance+TransmissionType+Exotic+Hatchback,data=train)
summary(forward.regression)
# This model has higher adjusted R-squared compared to the Log transformed model
par(mfrow=c(2,2))
plot(forward.regression)
library(olsrr)
ols_plot_cooksd_bar(forward.regression)
ols_plot_resid_stand(forward.regression)
# Cook's D plot looks okay - Much better than Log Transformed Plot
#Observation 3748,9152,9153 seem to have problem on the residual plot
summary(train[3748,]) # Ford Focus ... Not sure if we can delete this...
summary(train[9152,]) # Bugatti ... Delete?
summary(train[9153,]) # Bugatti ... Delete?
summary(train[1043,]) # Volvo ... Not seem to cause too much trouble on Cook's D

########################################
# Try Backward elimination model.#######
########################################
reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=966)
par(mfrow=c(1,3))
bics<-summary(reg.bwd)$bic
plot(1:967,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.bwd)$adjr2
plot(1:967,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.bwd)$rss
plot(1:967,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.bwd,248)
# Predictor for backward elimination method based on lowest BIC value
# Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic
backward.regression = lm(MSRP~Make+Model+EngineFuelType+VehicleSize+Crossover+HighPerformance+Hybrid+VehicleStyle+Exotic,data=train)
summary(backward.regression)
# Forward select has higher adjusted R-squared compared to backward method
par(mfrow=c(2,2))
plot(backward.regression)

@@

#Dummy code test set to see performance compare to other methods
xtest<-model.matrix(MSRP~.,test)[,-1]
ytest<-test$MSRP #36461071

@@

#####################################################
# Compare Forward vs. Backward on Test set to LASSO #
#####################################################
# Test result
test.result = test$MSRP
# str(test)
test = test[,-15] #delete MSRP for prediction
#Forward test results
forward.pred = predict(forward.regression,test)
# See forward test ASE
testMSE_Forward<-mean((test.result-forward.pred)^2)
testMSE_Forward #46312425

#Backward test results
backward.pred = predict(backward.regression,test)
# See backward test ASE
testMSE_Backward<-mean((test.result-backward.pred)^2)
testMSE_Backward #48317165

>>>>

```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
# You need to adjust the work environment to run this code
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project1")
# Directly import cleaned dataset and start from here
CarData = read.csv("CleanedCarData.csv")
###################################################################################
# Restrict prediction range to over cars under 1 million... To improve precision###
###################################################################################
CarData = CarData %>% filter(MSRP < 1000000)
# Convert factor variables
variables.to.factor = c("Make","Model","EngineFuelType","TransmissionType","DrivenWheels","MarketCategory","VehicleSize","VehicleStyle","Crossover","Diesel","Exotic","Luxury","HighPerformance","FactoryTuner","Performance","FlexFuel","Hatchback","Hybrid","N_A")
CarData[variables.to.factor] = lapply(CarData[variables.to.factor],factor)
summary(CarData)
# Plot agsint numeric variables
# CarData %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity) %>% ggpairs()
CarData = CarData[,-10] #delete MarketCategory
# Try different way of sampling by Models

######### Try Something else to split data ############
# First 836 are Models that only have 5 observations or less - deleted 6 card over 1 million dollars
ModelList = CarData%>% group_by(Model) %>% mutate(ModelCount = n()) %>% arrange(ModelCount)
#arrange(ModelList,desc(count))%>%print(n = Inf)
# Can only split train/test/validate set on Models that has 6 observations or more
ModelList = ModelList[,-27]
Not.Able.To.Split = ModelList[(1:836),]
Able.To.Split = ModelList[-(1:836),]

@@

# Joining train set with dataset that been left out
train = bind_rows(train,Not.Able.To.Split)

dim(train)

str(train)
str(CarData)
# Use Forward variable selection method to select possible variables as predictors
########################################
# Try Forward selection model.##########
########################################
library(leaps)
reg.fwd=regsubsets(MSRP~.,data=train,method="forward",nvmax=963)
par(mfrow=c(1,3))
bics<-summary(reg.fwd)$bic
plot(1:964,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.fwd)$adjr2
plot(1:964,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.fwd)$rss
plot(1:964,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.fwd,349)
# By looking at this model, it has selected following predictors:
# Make+Model+Year+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+TransmissionType+Exotic+Hatchback
forward.regression = lm(MSRP~Make+Model+Year+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+TransmissionType+Exotic+Hatchback,data=train)
summary(forward.regression)
# This model has higher adjusted R-squared compared to the Log transformed model
par(mfrow=c(2,2))
plot(forward.regression)
library(olsrr)
ols_plot_cooksd_bar(forward.regression)
ols_plot_resid_stand(forward.regression)
# Cook's D plot looks okay 
# Residual plot... ???

########################################
# Try Backward elimination model.#######
########################################
reg.bwd=regsubsets(MSRP~.,data=train,method="backward",nvmax=963)
par(mfrow=c(1,3))
bics<-summary(reg.bwd)$bic
plot(1:964,bics,type="l",ylab="BIC",xlab="# of predictors")
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(reg.bwd)$adjr2
plot(1:964,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(reg.bwd)$rss
plot(1:964,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

# how many predictors gives the best BIC?
which(bics==min(bics))
coef(reg.bwd,410)
# Predictor for backward elimination method based on lowest BIC value
# Make+Model+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+FactoryTuner+FlexFuel+Exotic+Hatchback
backward.regression = lm(MSRP~Make+Model+EngineFuelType+NumberOfDoors+VehicleSize+VehicleStyle+CityMPG+Luxury+FactoryTuner+FlexFuel+Exotic+Hatchback,data=train)
summary(backward.regression)
# Forward select has higher adjusted R-squared compared to backward method
par(mfrow=c(2,2))
plot(backward.regression)

@@

#Dummy code test set to see performance compare to other methods
xtest<-model.matrix(MSRP~.,test)[,-1]
ytest<-test$MSRP #36302210

@@

#####################################################
# Compare Forward vs. Backward on Test set to LASSO #
#####################################################
# Test result
test.result = test$MSRP
# str(test)
test = test[,-15] #delete MSRP for prediction
#Forward test results
forward.pred = predict(forward.regression,test)
# See forward test ASE
testMSE_Forward<-mean((test.result-forward.pred)^2)
testMSE_Forward #46054176

#Backward test results
backward.pred = predict(backward.regression,test)
# See backward test ASE
testMSE_Backward<-mean((test.result-backward.pred)^2)
testMSE_Backward #44935730
